Sparse Function-space Representation \\ of Neural Networks

### 2023-05-12 paper/main.tex

\textbf{Neural Network} In supervised learning, given a data set $\dataset = \{(\mathbf{x}_{i} , \mathbf{y}_{i})\}_{i=1}^{n}$, with $\mathbf{x}_i \in \inputDomain$ and $\mathbf{y}_i \in \outputDomain$, the weights $\weights \in \R^{P}$ of a neural network
, $f_\mathbf{w} : \inputDomain \mapsto \outputDomain$, 
are usually trained to minimize the (regularized) empirical risk,
\begin{align} \label{eq-empirical-risk}
  \weights^{*} = 
  \argmin_{\weights} \mathcal{L}(\dataset,\weights) =
  \argmin_{\weights} \sum_{i=1}^{n} \ell(f_\weights(\mathbf{x}_{i}), \mathbf{y}_i) + \delta \mathcal{R}(\weights).
\end{align}
If $\ell(f_\weights(\mathbf{x}_{i}), \mathbf{y}_i) = -\log(p(\mathbf{y} \mid f_\weights(\mathbf{x}_{i}))$ and $\mathcal{R}(\weights) = \|\weights\|^{2}_2$ then we can view \cref{eq-empirical-risk} as the MAP solution to a Bayesian objective, where $p(\vw) = \Norm(\vzeros, \delta^{-1} \MI)$.

\textbf{Neural network to GP prior} Our goal is to find a kernel from which to build a sparse Gaussian process model, central to this is a kernel function  $\kappa(\vx,\vx')$, where $\vx, \vx'$ are input vectors in $\R^{\mathrm{D}}$. If we had the following function parameterization $g(\vx) = \phi(\vx)^\top \vw$ where $\phi(\vx)$ is a feature map, we can convert from weight space prior to function space prior giving the following kernel $\kappa(\vx, \vx) = \delta^{-1} \phi(\vx) \phi(\vx)^\T$ see \citep{rasmussen2006gaussian} for details. We can approximate the feature map of the neural network  $f_\weights(\mathbf{x}_{i})$  by using the Jacobian at $\weights^*$ thus our GP prior becomes:
\begin{align} \label{eq-laplace-approx-function-space}
 g(\vx_i) \sim \mathcal{N} \left( \mu(\vx_i), k(\vx_i, \vx_i) \right) \quad
  \mu(\vx)
  = \vzeros \quad
  k(\mathbf{x}, \mathbf{x}')
  = \frac{1}{\delta} \Jac{\weights^*}{\mathbf{x}}\Jac{\weights^*}{\mathbf{x}'}^\top,
\end{align}
where the kernel is the neural tangent kernel (NTK) ,with Jacobian given by $\Jac{\weights}{\mathbf{x}} := \nabla_\weights f(\mathbf{x})^\top \in \R^{C \times P}$. Alternatively we could have followed the derivation in \citet{immerImprovingPredictionsBayesian2021} where we perform a local linearization of the neural network prediction function $f^{\textrm{lin}}_{\vw^*}(\vx; \vw) = f_{\vw^*}(\vx) + \Jac{\weights^*}{\mathbf{x}} (\vw - \vw^*)$. The difference with our GP prior mean function and \cref{eq-laplace-approx-function-space} is a different mean prior function which comes from $f^{\textrm{lin}}_{\vw^*}(\vx; \vzeros) = f_{\vw^*}(\vx) - \Jac{\weights^*}{\mathbf{x}} \vw^*$, essentially they use the difference from the linear model approximation and the neural network function, which does not seem useful as a prior mean function, so instead we use $ \mu(\vx_i) = \vzeros$ as is common in GP literature and derived from a feature map conversion.

@@

Replacing \cref{eq:dual_param_laplace} in to \cref{eq-laplace-approx-function-space} and we arrive at our GP model based on the converged neural network.

\textbf{Dual Gaussian Process} A Gaussian process characterizes a distribution over functions in terms of the first two moment.

A Gaussian process (GP) is a prior over functions $f \sim \GP(\mathbf{0}, \kappa(\cdot, \cdot))$. The prior is characterized by a covariance kernel function $\kappa(\vx,\vx')$, where $\vx, \vx'$ are input vectors in $\R^{\mathrm{D}}$. Once combined with a data set ${\mathcal{D} = \{(\vx_i, y_i)\}_{i=1}^n}$ with $n$ examples, each modelled by the likelihood $p(\vy_i \mid \vf_i)$ at function $\vf_i = f(\vx_i)$\, we can form a posterior proccess using multivariate conditioning. The drawbacks of GPs reside in their lack of scaling to large datasets and non-tractability with non-Gaussian likelihoods.

>>>>

\textbf{Neural Network} In supervised learning, given a data set $\dataset = \{(\mathbf{x}_{i} , \mathbf{y}_{i})\}_{i=1}^{n}$, with $\mathbf{x}_i \in \inputDomain$ and $\mathbf{y}_i \in \outputDomain$, the weights $\weights \in \R^{P}$ of a neural network
, $f_\mathbf{w} : \inputDomain \mapsto \outputDomain$, 
are usually trained to minimize the (regularized) empirical risk,
\begin{align} \label{eq-empirical-risk}
  \weights^{*} = 
  \argmin_{\weights} \mathcal{L}(\dataset,\weights) =
  \argmin_{\weights} \sum_{i=1}^{n} \ell(f_\weights(\mathbf{x}_{i}), \mathbf{y}_i) + \delta \mathcal{R}(\weights).
\end{align}
If $\ell(f_\weights(\mathbf{x}_{i}), \mathbf{y}_i) = -\log(p(\mathbf{y} \mid f_\weights(\mathbf{x}_{i}))$ and $\mathcal{R}(\weights) = \frac{1}{2}\|\weights\|^{2}_2$ then we can view \cref{eq-empirical-risk} as the MAP solution to a Bayesian objective, where $p(\vw) = \Norm(\vzeros, \delta^{-1} \MI)$.

\textbf{Neural network to GP prior} A Gaussian process characterizes a distribution over functions in terms of the first two moment. The key quantity of a Gaussian process is the prior covariance function $\kappa(\vx,\vx')$, where $\vx, \vx'$ are input vectors in $\R^{\mathrm{D}}$. Our goal is to find a kernel based on the neural network from which to build Gaussian process posterior. If we had the following function parameterization $g(\vx) = \phi(\vx)^\top \vw$ where $\phi(\vx)$ is a feature map, we can convert from weight space prior to function space prior giving the following kernel $\kappa(\vx, \vx) = \delta^{-1} \phi(\vx) \phi(\vx)^\T$ see \citep{rasmussen2006gaussian} for details. We can approximate the feature map of the neural network  $f_\weights(\mathbf{x}_{i})$  by using the Jacobian at $\weights^*$ thus our GP prior becomes:
\begin{align} \label{eq-laplace-approx-function-space}
 g(\vx_i) \sim \mathcal{N} \left( \mu(\vx_i), k(\vx_i, \vx_i) \right) \quad
  \mu(\vx)
  = \vzeros \quad
  k(\mathbf{x}, \mathbf{x}')
  = \frac{1}{\delta} \Jac{\weights^*}{\mathbf{x}}\Jac{\weights^*}{\mathbf{x}'}^\top,
\end{align}
where the kernel is the neural tangent kernel (NTK) ,with Jacobian given by $\Jac{\weights}{\mathbf{x}} := \nabla_\weights f(\mathbf{x})^\top \in \R^{C \times P}$. Alternatively we could have followed the derivation in \citet{immerImprovingPredictionsBayesian2021} where we perform a local linearization of the neural network prediction function $f^{\textrm{lin}}_{\vw^*}(\vx; \vw) = f_{\vw^*}(\vx) + \Jac{\weights^*}{\mathbf{x}} (\vw - \vw^*)$. The difference with our GP prior mean function and \cref{eq-laplace-approx-function-space} is a different mean prior function which comes from $f^{\textrm{lin}}_{\vw^*}(\vx; \vzeros) = f_{\vw^*}(\vx) - \Jac{\weights^*}{\mathbf{x}} \vw^*$, essentially they use the difference from the linear model approximation and the neural network function, which does not seem useful as a prior mean function, so instead we use $ \mu(\vx_i) = \vzeros$ as is common in GP literature and derived from a feature map conversion.

@@

Replacing \cref{eq:dual_param_laplace} in to \cref{eq-laplace-approx-function-space} and we arrive at our GP model based on the converged neural network. Again this is similar to what was arrived at for \citet{immerImprovingPredictionsBayesian2021} for the posterior variance function, but they instead use the $f_{\vw^*}(\vx)$ for the posterior mean and fail to spot the significance of the dual variables. The problem with \cref{eq:gp_pred} is that to make predictions and compute variances we must incur a cost of $O(n^3)$ which limits the use of the GP on large datasets.

\textbf{Sparse Dual Gaussian Process} Sparse Gaussian processes are a way to reduce the computational complexity of a full GP using a low rank approximation many approaches have been devised see \citet{quinonero2005unifying} for an overview. Given we have our dual variables derived from our neural network predictions and kernel function we could make use any approximation. We decide to follow the the \citep{titsias2009variational}, which is to say that our marginal distribution $q_{\vu}(f_i)  = \int p(f_i  \mid \vu) q(\vu)  d\vu$
